---
title: "additional_functions"
output: html_notebook
---
```{r}
install.packages("tidyr")
install.packages("xts")
```

tidyr

In finance, you will often download data from Bloomberg or Yahoo Finance where each column is a different asset, and each row is a date. This is called wide data.

However, packages like dplyr and ggplot2 strongly prefer long data—where every row is a single observation. The tidyr package provides two  functions to flip your data back and forth: pivot_longer() and pivot_wider().

```{r}
library(tidyr)
library(dplyr)

# 1. The Wide Data (Typical finance download format)
wide_prices <- data.frame(
  Date = c("2024-01-01", "2024-01-02"),
  AAPL = c(180.5, 181.2),
  MSFT = c(330.1, 332.5)
)
wide_prices
# 2. Making it Long
long_prices <- wide_prices %>%
  pivot_longer(
    cols = c(AAPL, MSFT),      # The columns we want to melt down
    names_to = "Ticker",       # The new column name for the column headers
    values_to = "Price"        # The new column name for the actual numbers
  )

long_prices
```
You have downloaded a small sample of government bond yields across different maturities. The data is currently "wide," and you need it "long" so you can eventually plot a yield curve.

```{r}
bond_yields_wide <- data.frame(
  Date = c("2024-03-01", "2024-03-02"),
  Yield_2Yr = c(4.55, 4.58),
  Yield_10Yr = c(4.20, 4.25),
  Yield_30Yr = c(4.35, 4.38)
)
bond_yields_wide
bond_yields_wide %>% pivot_longer(cols = colnames(bond_yields_wide)[2:length(colnames(bond_yields_wide))], names_to = "Maturity Period", values_to = "Yield %")

# improvement
bond_yields_long<-bond_yields_wide %>% pivot_longer(cols = -Date, names_to = "Maturity_Period", values_to = "Yield_Pct")
bond_yields_long
```

Time Series Handling

Why xts is vital for finance:

Date/Time Indexing: The row names are the timestamps.
Elegant Subsetting: You can query time ranges using natural strings like "2024-01" to get all of January, or "2024-01-01/2024-06-30" to get the first half of the year.
Alignment: If you merge two xts objects with different timestamps (e.g., a stock trading in the US and one in the UK with different holidays), xts automatically aligns them chronologically and handles the missing days perfectly.

```{r}
# Load the package (install.packages("xts") if needed)
library(xts)

# 1. Raw data
prices <- c(150, 152, 151, 155, 158)
dates <- as.Date(c("2024-04-01", "2024-04-02", "2024-04-03", "2024-04-04", "2024-04-05"))

# 2. Create the xts object
aapl_ts <- xts(x = prices, order.by = dates)

# 3. Time-based subsetting: Give me data from April 3rd onwards
print(aapl_ts["2024-04-03/"])
```

```{r}
btc_prices <- c(68000, 69500, 69200, 71000, 70500, 72000, 71500)
btc_dates <- as.Date(c("2024-03-11", "2024-03-12", "2024-03-13", 
                       "2024-03-14", "2024-03-15", "2024-03-16", "2024-03-17"))
```

```{r}
btc_ts <- xts(x = btc_prices, order.by = btc_dates)
btc_ts["2024-03-16/2024-03-17"]
```
Linear Modelling
The lm() function stands for "linear model." It fits a line of best fit through your data using Ordinary Least Squares (OLS) regression.
```{r}
# Mock data
returns_data <- data.frame(
  Stock_Ret = c(0.012, -0.005, 0.021, -0.011, 0.008),
  Market_Ret = c(0.010, -0.008, 0.015, -0.015, 0.005)
)

# Run the regression (Stock Returns as a function of Market Returns)
capm_model <- lm(Stock_Ret ~ Market_Ret, data = returns_data)

# Print the statistical summary (coefficients, p-values, R-squared)
summary(capm_model)
```

```{r}
algo_data <- data.frame(
  Volume_Millions = c(1.2, 2.5, 0.8, 3.1, 1.5, 4.0),
  Volatility_Pct = c(1.1, 2.3, 0.9, 2.8, 1.4, 3.5)
)

vol_model <- lm(Volatility_Pct ~ Volume_Millions, data = algo_data)

summary(vol_model)
```

