---
title: "test"
output: html_notebook
---

Vectors 
- In R, there is no such thing as a "scalar" (a single number). Even a single number is just a vector of length 1
- We use the c() function (which stands for combine or concatenate) to create vectors. 
- R excels at vectorized math, meaning you can perform operations on whole vectors without writing for loops.

```{r}
# Creating a numeric vector of stock prices
prices <- c(150.5, 152.0, 149.5, 155.2)

# Vectorized operations: Calculating simple returns against a base price
base_price <- 150.0
returns <- (prices / base_price) - 1

# Printing the result
print(returns)
```

By default, R fills matrices column by column. If you want it to fill row by row, you must specify byrow = TRUE

```{r}
# Creating a 2x2 matrix of portfolio weights
portfolio_weights <- matrix(c(0.4, 0.6, 0.5, 0.5), nrow = 2, ncol = 2)
portfolio_weights_by_row_fill <- matrix(c(0.4, 0.6, 0.5, 0.5), nrow = 2, ncol = 2, byrow = TRUE)

# Assigning names to rows and columns for clarity
rownames(portfolio_weights) <- c("Asset A", "Asset B")
colnames(portfolio_weights) <- c("Portfolio 1", "Portfolio 2")

print(portfolio_weights)
print(portfolio_weights_by_row_fill)
```

DataFrames are the equivalent of a spreadsheet or a SQL table. 
- Unlike a matrix, each column in a DataFrame can be a different data type (e.g., numeric prices in one column, character tickers in another).

```{r}
market_data <- data.frame(
  Ticker = c("AAPL", "GOOGL", "MSFT"),
  Price = c(180.5, 140.2, 330.1),
  TechSector = c(TRUE, TRUE, TRUE)
)
market_data
```

Additional context
<- is for Environment Assignment: You use the arrow when you are creating a brand new object (like a variable, vector, or dataframe) and saving it into your workspace/memory.

= is for Argument Assignment: You use the equals sign inside the parentheses of a function to pass values to specific arguments or to name things within that function.

Task

Your Tasks:
Vectors: Create two numeric vectors.
- stock_a with prices: 100, 102, 101
- stock_b with prices: 50, 55, 53

Matrix: Combine these two vectors into a single 3x2 matrix representing your price data. (Hint: You can use the matrix() function we discussed, or look up a handy R function called cbind() which binds columns together).

DataFrame: Create a DataFrame that includes three columns:
- Day: A vector containing the days 1, 2, and 3.
- Stock_A: Your first price vector.
- Stock_B: Your second price vector.

```{r}
stock_a <- c(100,102,101)
stock_b <- c(50,55,53)
matrix_ver <- matrix(c(stock_a,stock_b), nrow = 3, ncol = 2)
cbind_matrix_ver <- cbind(stock_a, stock_b)
rownames(matrix_ver) <- c(1,2,3)
colnames(matrix_ver) <-c("Stock_A","Stock_B")
matrix_ver
```
```{r}
 data.frame(
  Ticker = c(1,2,3),
  Stock_A = c(100,102,101),
  Stock_B = c(50,55,53))
```

