---
title: "data_wrangling"
output: html_notebook
---
```{r}
install.packages("dplyr")
```


Key: pipe (%>% or the newer base R version |> ).
- It takes whatever is on its left and feeds it into the function on its right. This allows you to chain commands together beautifully without nesting functions inside each other.

1. select(): Working with Columns
Use select() when you want to keep, drop, or reorder specific columns in your DataFrame.

2. filter(): Working with Rows
Use filter() to subset your data based on logical conditions (e.g., "only show me rows where the price is greater than 100"). 
- Note: R uses == for logical equals.

3. mutate(): Creating New Data
Use mutate() to add brand new columns to your DataFrame, usually calculated from existing columns.

```{r}
library(dplyr)

# 1. Our raw DataFrame
raw_portfolio <- data.frame(
  Ticker = c("AAPL", "GOOGL", "TSLA", "MSFT", "AMZN"),
  Sector = c("Tech", "Tech", "Auto", "Tech", "Consumer"),
  Price = c(185.0, 140.0, 250.0, 335.0, 130.0),
  Shares = c(10, 15, 5, 20, 50)
)

# 2. The dplyr pipeline
tech_summary <- raw_portfolio %>%
  filter(Sector == "Tech") %>%                  # Keep ONLY Tech stocks
  select(Ticker, Price, Shares) %>%             # Drop the Sector column
  mutate(Position_Value = Price * Shares)       # Calculate total $ value

print(tech_summary)
```

```{r}
fx_data <- data.frame(
  Pair = c("EUR/USD", "GBP/USD", "JPY/USD", "EUR/USD", "GBP/USD", "JPY/USD"),
  Day = c(1, 1, 1, 2, 2, 2),
  Open_Rate = c(1.08, 1.26, 0.0068, 1.09, 1.25, 0.0069),
  Close_Rate = c(1.09, 1.27, 0.0067, 1.08, 1.26, 0.0070),
  Volume = c(1000, 850, 1200, 1100, 900, 1300)
)
```

Your Tasks:
Using dplyr and the pipe (%>%):

Filter the fx_data to only show trades that happened on Day 1.

Select only the Pair, Open_Rate, and Close_Rate columns (drop Day and Volume).

Mutate a new column called Daily_Change which is calculated as Close_Rate - Open_Rate.

```{r}
fx_data %>% filter(Day == 1) %>% select(Pair, Open_Rate, Close_Rate) %>% mutate(Daily_Change = Close_Rate - Open_Rate)
```


