---
title: "data_viz"
output: html_notebook
---


```{r}
install.packages("ggplot2")

```
```{r}
fx_data <- data.frame(
  Pair = c("EUR/USD", "GBP/USD", "JPY/USD", "EUR/USD", "GBP/USD", "JPY/USD"),
  Day = c(1, 1, 1, 2, 2, 2),
  Open_Rate = c(1.08, 1.26, 0.0068, 1.09, 1.25, 0.0069),
  Close_Rate = c(1.09, 1.27, 0.0067, 1.08, 1.26, 0.0070),
  Volume = c(1000, 850, 1200, 1100, 900, 1300)
)
```

Here are the three essential layers every ggplot needs:

Data: The DataFrame you are plotting (e.g., fx_data).
Aesthetics (aes): How your data maps to visual elements like the x-axis, y-axis, color, or size.
Geometries (geom): The actual shape of the plot (e.g., geom_point() for a scatter plot, geom_line() for a line chart, geom_col() for a bar chart).

```{r}
# Load the package
library(ggplot2)

# Building the plot layer by layer
ggplot(data = fx_data, aes(x = Open_Rate, y = Close_Rate, color = Pair)) +
  geom_point(size = 3) +                            # Adds the scatter plot dots
  labs(title = "FX Open vs Close Rates",            # Adds titles and labels
       x = "Opening Price", 
       y = "Closing Price") +
  theme_minimal()                                   # Gives it a clean, professional look
```
Using ggplot2, write the code to create a bar chart (column chart):

Data: Use fx_data.
Aesthetics (aes): Map the x-axis to Pair and the y-axis to Volume.
Geometry: Add the appropriate geom for a column/bar chart. (Hint: Look up geom_col()).
Add a title using the labs() layer!

```{r}
ggplot(data = fx_data, aes(x= Pair, y=Volume)) +
         geom_col() +
       labs(title = "FX Pair vs Volume") + theme_minimal()
```


